Extracting sunbeams from cucumbers: Computational reproducibility is essential for archaeological science

### 2025-02-18 analysis/paper/paper.qmd
writing conclusion

In the classic satirical novel *Gulliver's Travels* (1726) by Irish writer Jonathon Swift Gulliver visits the fictional Grand Academy of Lagado in Balnibarbi, a caricature of the Royal Society of London, and meets several researchers working on wildly impractical projects, including one attempting to extracting sunbeams out of cucumbers. This is usually interpreted as a subversive and anti-colonial parody depicting institutionalized research as an absurd fund-raising activity with no practical benefits to society [@alff2014swift; @nicolson1937scientific]. @feinbergExtractingSunbeamsCucumbers2011

I have presented a bibliometric analysis on the status of archaeology as a science, showing distinct disunity, but that we generally behave as social scientists. If archaeological science is to continue to progress through increased consensus through the accumulation of reliable facts and methods, it is essential for researchers to take computational reproducibility seriously. Computers have become a central field and laboratory instrument for so much of our work, so we have an ethical duty to document how we change our data as it flows through silicon just as carefully as we document the operating parameters of a mass spectrometer or any other field or laboratory instrument. However, the current state of quantitative archaeology, with most researchers not using open source code, is comparable to the secrecy of alchemy prior to the emergence of chemistry. Abandoning this habit of secrecy in favour of reproducibility is vital if we are to avoid a future where our journals are filled with pretty pictures that the reader has no hope of repeating or adapting in their own work. It is time for computational reproducibility to be considered a minimum requirement for evaluating the integrity and usefulness of empirical results.

>>>>

In the classic satirical novel *Gulliver's Travels* (1726) by Irish writer Jonathon Swift Gulliver visits the fictional Grand Academy of Lagado in Balnibarbi, a caricature of the Royal Society of London, and meets several researchers working on wildly impractical projects, including one attempting to extracting sunbeams out of cucumbers. This is usually interpreted as a subversive and anti-colonial parody depicting institutionalized research as an absurd fund-raising activity with no practical benefits to society [@alff2014swift; @nicolson1937scientific]. It has also been used as a metaphor for the difficulty of extracting useful data out of published  @feinbergExtractingSunbeamsCucumbers2011

I have presented a bibliometric analysis on the status of archaeology as a science, showing distinct disunity that is increasing over time,. On average we generally behave as social scientists, with some elements in common with harder sciences. If archaeological science is to continue to progress through increased consensus through the accumulation of reliable facts and methods, it is essential for researchers to take computational reproducibility seriously. Computers have become a central field and laboratory instrument for so much of our work, so we have an ethical duty to document how we change our data as it flows through silicon just as carefully as we document the operating parameters of a mass spectrometer or any other field or laboratory instrument. However, the current state of quantitative archaeology, with most researchers not using open source code, is comparable to the secrecy of alchemy prior to the emergence of chemistry. Abandoning this habit of secrecy in favour of reproducibility is vital if we are to avoid a future where our journals are filled with pretty pictures that the reader has no hope of repeating or adapting in their own work. It is time for computational reproducibility to be considered a minimum requirement for evaluating the integrity and usefulness of empirical results.
